% vim: fo=aw2tq tw=100 spell
\chapter{Design and Implementation}

\section{Audio output}
\label{sec:design:audio_output}

% TODO: sound/music theory somewhere?
The output device was jointly designed with my lab partner.  The most important property of any 
output device developed is obviously the ability to output a waveform that would audibly represent 
the intended musical note (and instrument).  However, since the network data also includes volume 
information, this needed to be taken into account.  This could be achieved in two ways:
\begin{enumerate}
\item Implementing a volume scaling function in software
\item Scaling the volume with more programmable hardware
\end{enumerate}
The first option has the advantage of giving the simplest, and therefore easiest, hardware to 
design.  Unfortunately, scaling the volume of a waveform during playback would involve modifying the 
amplitude of every step of a waveform with multiplication.  This would result in playback being a 
very processor-intensive task, which is fairly impractical for the speed and simplicity of our CPU.  
On the other hand, implementing volume scaling in hardware removes this burden from the CPU, freeing 
up the processing time for other tasks.

We chose to implement the volume scaling in hardware, and gave the hardware a very simple interface 
--- an 8-bit ``waveform'' input and an 8-bit ``volume'' input.  This would later make outputting a 
waveform a very fast software task --- one I/O operation per sample in the waveform, and one I/O 
operation for every volume value received.

The conversion from the two 8-bit values to the final audio signal can be described as a two-stage 
digital-to-analogue converter (DAC) with a final amplification stage.  A simple circuit of a DAC and 
inverting operational amplifier will give an output voltage between ground and $-V_{ref}$ 
proportional to the digital input, where $V_{ref}$ is the reference voltage of the DAC.  (The 
maximum output in practice is about $-V_{ref}\times0.85$.)  We decided to use this concept to 
develop a two-stage DAC consisting of two such circuits, where the output of the first became the 
$V_{ref}$ of the second.  The first becomes the ``volume'' DAC, effectively pre-scaling any signal 
output by the second, which would receive the waveform sample values.  Note that this design does 
not prevent the use of software scaling if desired, but gives the attractive possibility of avoiding 
it.

To interface with the SBC, the DAC inputs were connected to ports A and B of the parallel I/O 
controller.  Though the network data contained 7-bit volume values, and the hardware needed 8-bit 
values, we decided not to ``left-shift'' the data in hardware (wiring parallel I/O data pins 0--6 to 
DAC data pins 1--7) to maintain flexibility of the hardware.  (In my solution the volume values are 
left-shifted in software.)  The result of this was that existing I/O ports on the SBC were used, 
removing the need to create extra hardware for chip selection based on I/O address, and the DAC 
operation was simple because the parallel I/O is latched (data stays on the outputs between writes).

There are a few things worth noting in the circuit diagram for the audio output device (appendix 
\ref{appendix:circuit_diagram}):

\subsection{DACs In ``Flow-Through'' Mode}

The DAC0832 used for this project is a double-buffered digital to analogue converter.  Single 
buffering is generally useful for holding the input value in the DAC between writes.  Double 
buffering adds the ability to defer the write until some other signal --- this is especially useful 
in cases where several DACs need to change simultaneously but their data is received asynchronously, 
since the write between the first and second buffers can be triggered on all DACs at the same time.

In this instance however, the ``single-buffered'' effect is achieved through the parallel I/O having 
latched outputs, and the ``double-buffered'' effect is unneeded as the volume and waveform data are 
not directly related and do not need to be applied simultaneously.  By hard-wiring all of the 
signals related to writing and buffering to their ``enable'' value, the output of the DAC follows 
the digital input directly.  This means that writes to the parallel I/O have the effect of changing 
the DAC values instantly.

\subsection{DAC Potential Divider}

In the top-right of the circuit diagram there is a potential divider circuit between the $V_{ref}$ 
input of the volume DAC and $+12V$ (R1, R2 and C2).  It supplies a $\sim7.1V$ potential, made very 
smooth by the $22\mu{}F$ capacitor --- the DAC is very sensitive to noise on $V_{ref}$.

\subsection{Feedback Capacitors}

The capacitors C1 and C5 --- between the output and inverting input of both operational amplifiers) 
--- are to remove ``ringing'' and ``overshoot'' from the output signals (as recommended by the 
DAC0832 datasheet\cite{dac0832}).  Initially both capacitors were $68pF$, sufficient to remove any 
noticeable overshoot from the signal.  However, once the hardware and software solutions were 
complete and the maximum sample rate was known, the value of C5 was increased to $10nF$ to smooth 
the output signal and make it sound less ``square'' --- a problem present due to the quick value 
changes compared to the time between changes.

\subsection{Audio Amplifier}

The audio amplifier (IC5) is set up in a similar way to that described in the LM386 audio amplifier 
datasheet\cite{lm386} (diagram titled ``Amplifier with Gain = 20, Minimum Parts'') --- we were not 
too interested in voltage gain, our reason for using an amplifier was to supply more current so that 
driving a speaker did not result in clipping of the waveform.

\subsection{Pre-Amplifier Potential Divider}

The datasheet for the LM386\cite{lm386} states that the input voltage for the audio amplifier is 
$\pm0.4V$ --- above this voltage the signal is clipped (as we found through experimentation).  
However, the maximum output amplitude of the second DAC stage is $\sim5.1V$ ($7.1\times0.85^2$), so 
this needed to be brought within the acceptable input range of the LM386.  This was achieved with a 
potential divider comprising of a $150K\Omega$ resistor (R3) and a $10K\Omega$ potentiometer (P1).  

The potentiometer was used to give a ``physical'' volume control to the system.  The $220\mu{}F$ 
capacitor (C6) before the potential divider AC-couples the signal, putting it in the range 
$\sim\pm2.55V$.  When the potentiometer is at it's maximum resistance, the signal into the LM386 is 
in the range $\sim\pm0.16V$.  This is actually lower than necessary, because when I calculated the 
resistor values I failed to take into account the AC coupling (I was aiming for a maximum voltage of 
$\pm0.32V$).  However, this is of little consequence since the maximum output can still cause the 
amplifier to clip.


\section{Keypad handler}
\label{sec:design:keypad_handler}

The keypad is an awkward device for trying to get a single value from --- it interrupts continuously 
for the duration of the key press.  When implementing the keypad handler, this was the main problem 
I had to overcome.

My solution was to create a kind of lockout for that particular interrupt handler.  The result is a 
variable that acts as an ``already handled'' flag for keypad interrupts.  When a keypad interrupt 
happens, the flag is checked --- if it's non-zero, then the interrupt gets effectively ignored.  If 
it's zero, the interrupt handler executes as normal.  At the end of the interrupt handler, there are 
a few NOP (no operation) instructions before setting the lockout variable back to zero and returning 
from the interrupt, but \emph{after} the interrupts are re-enabled.  The idea is that if the key is 
still pressed, it will interrupt during those NOPs, before the flag is cleared, and those interrupts 
will get ignored, but on the very last interrupt it will complete the handler.  This completely 
eliminates the possibility of accidental repeats.

For comparison, the alternative solution was to insert a delay after handling the keypress to allow 
time for the key to be released.  I found this method to be less than ideal, since if the delay is 
too long it can result in the system seeming unresponsive, and if too short then repeated interrupts 
can occur.  I felt it was better to maintain a direct response between action and reaction.

\section{Note lookup tables}
\label{notelookuptables}

Lookup tables are needed to contain the PRT value, divisor and note name (including octave) for 
every MIDI note.  The note names are in a separate lookup table to the other data as having an LCD 
information display was a lower priority than working sound playback and was implemented much later.

% TODO: expand on comparison of formats
The most efficient way to use a lookup value for a table is to multiply it by a power of two (left 
shifting), so blocks of data in the lookup table are also going to be $2^n$ bytes long.  However, 
the data for note playback is 3 bytes (a 16-bit PRT value plus an 8-bit divisor), so I was faced 
with the choice of having two lookup tables for the playback data or wasting $\frac{1}{4}$ of the 
space used by the table.  After writing the necessary assembler for reading from both possible 
formats and analysing their execution times, I found that the single table method resulted in faster 
code.  Combined with the fact I was not under pressure to conserve memory, I chose the single table 
method.

% TODO: expand on the result sorting
The Python script\footnote{"tools/pitchtable.py" in the submitted source code} I wrote to generate 
the lookup tables\footnote{"note\_lookup.s"} is based on the concept outlined in \ref{wavetables} 
for creating new frequencies from a single sample by modifying the sample rate and divisor.  PRT and 
divisor values are calculated for the frequency information in the pitch 
table\footnote{"tools/pitchtable.csv" - octave, MIDI note number, note name, frequency}.

The first step is calculating the PRT value for every divisor, $D$ from 1 to 255, first by 
calculating the sample rate using

\[R_{target} = \frac{F_{target} \times R_{source}}{F_{source} \times D}\]

(rearranged from a similar equation in \ref{wavetables}) and then converting this to a PRT value 
using

\[PRT = round\left(\frac{F_{CPU}}{D_{PRT}\times{}R_{target}}\right)\]

where $F_{CPU}$ is the CPU frequency (6,144,000) and $D_{PRT}$ is the PRT divisor (20).  Because the 
result has been rounded to the nearest integer, the effective frequency will be different to the 
desired frequency.  The effective frequency is calculated by using the $PRT$ and $R_{target}$ 
calculations in reverse:

\[F_{effective} = \frac{F_{source}\times{}D\times{}\frac{F_{CPU}}{D_{PRT}\times{}PRT}}{R_{source}}\]

The error, E is calculated with the following, giving a value in the range 0--1:

\[E = \frac{\left|F_{effective} - F_{target}\right|}{F_{target}}\]

% TODO: to joke or not to joke?
For each note, the results for all the divisors are collected together, discarding those where the 
sample rate is higher than 8kHz (a sensible limit given the CPU speed), and sorted in reverse order 
by $R^{1-E}$.  The idea is that the basic ordering aims for a high sample rate, but any kind of 
significant error greatly diminishes the value of the result.  The first result is taken from the 
sorted list, giving the PRT and divisor values the algorithm decides best represent the frequency.  
While almost seeming like mathematical voodoo, the worst error rate for any note in the resulting 
table is $1.13\%$, with an average sample rate of just under 7kHz.  Other sorting methods I tried 
had varied results, generally creating high sample rates with high error margins, or low error 
margins with bad sample rates.

The format of the resulting lookup table is:
\begin{nowordcount}
\begin{center}
\begin{tabular}{c | c | c | c | c}
Byte & 0 & 1 & 2 & 3 \\
\hline
Usage & PRT low & PRT high & Divisor & 00h \\
\end{tabular}
\end{center}
\end{nowordcount}

The script also produces a second table which contains a 4-character representation of the note 
created from the octave number and note name.
