% vim: fo=aw2tq tw=100 spell
\chapter{Introduction}

The aim of this project is to develop and implement an embedded system capable of synthesising one 
or more musical instruments, and playing these instruments in real-time based on a serial network 
data stream.

The base hardware supplied for this task is a Single-Board Computer (SBC) based on the Hitachi 64180 
CPU.  The 64180 is effectively an extended Z80, with several useful peripherals integrated, 
including Programmable Reload Timer (PRT), Direct Memory Access (DMA) and serial I/O controllers.  
It also includes a few extra instructions, mostly related to using the inbuilt peripherals, but 
notably including multiplication, and performance improvements over many Z80 instructions.  The CPU 
is clocked at 6.144MHz, and has a physical address space of 512KB, courtesy of an inbuilt Memory 
Management Unit (MMU), however it's logical address space is limited to 64KB, inheriting the Z80's 
use of 16-bit memory addresses.

In addition to the 64180 chip, many other useful peripherals are part of the SBC, including an 
I$^{2}$C controller, Real-Time Clock (RTC), 24-bit parallel digital I/O, 4-bit hexadecimal keypad 
and $16\times2$ character LCD text display.  The board is fitted with 96KB of RAM, and a ROM 
containing a very basic operating system that, among other things, supports loading software over a 
serial connection and starting execution at a specified memory location.

The audio output of the system developed should be suitable for use with an $8\Omega$-impedance 
speaker.  Additional hardware required to generate this output will be designed and built jointly 
with a lab partner.

% TODO: More info on the output
%Output is to be via a supplied $8\Omega$ impedance speaker.  Any other hardware required for the 
%output is to be designed and built jointly with a partner, since repeated changing of hardware on a 
%shared system for testing individual solutions is impractical, and the design of the hardware will 
%heavily influence the way in which the software is written.
