% vim: fo=aw2tq tw=100 spell
\chapter{Technical Specification}

\section{Operating Parameters}

\begin{itemize}
\item Power supply: $+12V$, $-12V$, $+5V$, Ground
\item Temperature: 0--70$^{\circ}$C\footnote{Minimum range suitable for all components.}
\end{itemize}

\section{Functionality}
\label{sec:spec:functionality}

\subsection{Audio Playback}

\begin{itemize}
\item Playable MIDI notes
    \begin{itemize}
    \item Maximum range: 0--127
    \item Optimal range: 0--106
    \item Accuracy (worst): $\pm1.13\%$
    \item Accuracy (avg.): $\pm0.5\%$
    \end{itemize}
\item Volume range
    \begin{itemize}
    \item 128 steps
    \item Linear scale
    \item Additional hardware volume control
    \end{itemize}
\item Supported instruments
    \begin{itemize}
    \item Bass guitar
    \item Cello
    \item Church Organ
    \item Piano
    \item Saxophone
    \item Melody
    \item Violin
    \item Trombone
    \item Trumpet
    \item French Horn
    \item Synth
    \item Electric Guitar
    \item Acoustic Guitar
    \item Flute
    \item Piccolo
    \end{itemize}
\end{itemize}

\subsection{Interface}
\begin{itemize}
\item Network stream
    \begin{itemize}
    \item 19,200bps baud rate
    \item 34-byte packet length
        \begin{itemize}
        \item 2 bytes per channel, in order --- pitch, volume
        \item Terminated by null and carriage return (00h, 0Dh)
        \end{itemize}
    \end{itemize}
\item 16-button keypad --- allows changing of channel at any time.
\item Terminal interface
    \begin{itemize}
    \item Change channel and instrument independently.
    \item Piano mode for manually playing notes with a keyboard --- layout based on an actual piano.
    \item Not necessary for operation --- system can be used standalone.
    \end{itemize}
\end{itemize}


\section{Cost}
\label{sec:spec:cost}

\begin{nowordcount}
\begin{center}
\begin{tabular}{c | c | r | r}
Part & Quantity & Per-unit & Total \\
\hline\hline
64180 SBC               & 1     & \pounds100.00 & \pounds100.00 \\
DAC0832LCN              & 2     & \pounds1.66 & \pounds3.32 \\
LM356N                  & 2     & \pounds0.40 & \pounds0.80 \\
LF386N-1                & 1     & \pounds0.26 & \pounds0.26 \\
$8\Omega$ speaker       & 1     & \pounds2.62 & \pounds2.62 \\
$10K\Omega$ potentiometer & 1     & \pounds0.72 & \pounds0.72 \\
Metal film resistor ($\pm1\%$)    & 3     & \pounds0.06 & \pounds0.18 \\
Electrolytic capacitor  & 5     & \pounds0.08 & \pounds0.40 \\
Ceramic capacitor       & 5     & \pounds0.07 & \pounds0.35 \\
PCB manufacture (40mm $\times$ 40mm) & 1     & \pounds2.90 & \pounds2.90 \\
\hline
\multicolumn{3}{r |}{\textbf{Total}} & \textbf{\pounds111.15} \\
\end{tabular}
\end{center}
\end{nowordcount}

Component prices (except for SBC) quoted from Farnell UK.  PCB manufacture quoted from PCBTrain 
quote calculator\footnote{\url{http://www.pcbtrain.co.uk/onestepquote.php}}.  Prices correct as of 
07/04/2008.  Based on a batch of 100 units.
